---
output: rmarkdown::github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r echo=FALSE, message=FALSE}
knitr::opts_chunk$set(message=FALSE, comment="")
```

# The Essentials

This is meant to serve as a very brief introduction to adopting tidyverse packages for data manipulation in R. It is just to give you a sense of the functionality of commonly used packages and hopefully convince you to take a deep dive in all that tidyverse has to offer. He we cover some essential functions that you will use on a daily basis as well as some useful R tricks and tips.

```{r}
library(magrittr)
library(dplyr)
library(tidyr)
library(purrr)
library(janitor)
library(glue)
library(reshape2)
```

Or just use `tidyverse`
```{r eval=FALSE}
library(tidyverse)
```

A fun penguins dataset for examples
```{r}
library(palmerpenguins)
df <- penguins
```

## Data Cleaning
```{r}
# What if the names were ugly?
colnames(df) <- c("Species", "ISLAND", "Bill_Length$(MM)", "bill.depth.mm", "flipper.length.mm", "body.mass.(g)", "SEX", "YEAR")
head(df)

df <- janitor::clean_names(df)
head(df)
```

```{r}
summary(penguins)
```

## `magrittr` 

**`%>%`**

Here we want to find the average bill length in inches.

```{r}
# Fine
mean_bill_length_in <- mean(df$bill_length_mm * 0.0393701, na.rm=TRUE)
print(mean_bill_length_in)

# Better
mean_bill_length_in <- df %>%
  dplyr::pull(bill_length_mm) %>%
  magrittr::multiply_by(0.0393701) %>%
  mean(na.rm=TRUE)

print(mean_bill_length_in)
```

**`set_colnames()`**
```{r}
df %>%
magrittr::set_colnames(paste0("Column_", seq(ncol(.)))) %>%
head()
```

You can also insert short hand functions!

```{r}
df %>%
dplyr::pull(bill_length_mm) %>%
(function(x) {
  return(x * 0.0393701)
}) %>%
mean(na.rm=TRUE)
```

## `dplyr`

Please check out the [full reference](https://dplyr.tidyverse.org/reference/index.html) for more functions

**`mutate()`**

Modify a new or existing column

```{r}
df %>%
dplyr::mutate(bill_length_in = bill_length_mm * 0.0393701) %>%
head()
```

**`filter()`**

Filter rows

```{r}
df %>%
dplyr::filter(species == "Chinstrap") %>%
head()
```

**`select()`**

Select columns

```{r}
df %>%
dplyr::select(species, body_mass_g) %>%
head()
```

**`pull()`**

Select columns and flatten to vector

```{r}
df %>%
dplyr::pull(body_mass_g) %>%
head()
```

**`arrange()`**

Order rows

```{r}
df %>%
dplyr::arrange(desc(body_mass_g)) %>%
head()
```

**`across()`**

Apply function across columns

```{r}
df %>%
dplyr::mutate(dplyr::across(c(bill_length_mm, bill_depth_mm), round)) %>%
head()
```

**`group_by()` + `summarise()`**
Group by columns and compute summary statistics

```{r}
df %>%
dplyr::group_by(species) %>%
dplyr::summarise(n=n(), mass=mean(body_mass_g, na.rm=TRUE))
```

**Conditional mutate**
```{r}
df %>%
dplyr::mutate(body_size=
      dplyr::case_when(body_mass_g > 5000 ~ "Massive",
                       body_mass_g > 4000 ~ "Big",
                       TRUE ~ "Small")) %>%
dplyr::pull(body_size) %>%
table()
```

## `purrr`

**Mutate with two or more columns**
```{r}
df.xyz <- data.frame(x=c(1,3), y=c(1,3), z=c(1,3))
head(df.xyz)
df.xyz <- df.xyz %>%
      dplyr::mutate(xyz = purrr::pmap(list(x, y, z), function(a, b, c) {
          a + b + c
      }))

head(df.xyz)
```

## `glue`

String manipulation and formatting

```{r}
name <- "Fred"
glue('My name is {name}.')
```

```{r}
df %>%
dplyr::select(species, bill_length_mm, sex) %>%
tidyr::drop_na() %>%
dplyr::mutate(desc = glue("This {sex} penguin ({species}) has a bill length of {bill_length_mm} mm")) %>%
head()
```

## `reshape2`

**`melt()`**

Melt a data frame for visualization

```{r}
df %>%
dplyr::select(species, bill_length_mm, bill_depth_mm) %>%
reshape2::melt(id=c("species")) %>%
head(5)
```

## Extra helpers

**Not in**
```{r eval=FALSE}
`%not_in%` <- Negate(`%in%`)
```

**Not null**
```{r eval=FALSE}
not_null <- Negate(is.null)
```

**Not NA**
```{r eval=FALSE}
not_na <- Negate(is.na)
```

**Remove nulls from a vector**
```{r eval=FALSE}
drop_nulls <- function(x){
  x[!sapply(x, is.null)]
}
```

**Iterate through all files in directory**
```{r eval=FALSE}
files <- list.files(path=path.to.dir, pattern="*.pattern", full.names=TRUE, recursive=FALSE)
lapply(files, function(fn) {
  do.something(fn)
})
```

**Apply function to a list**
```{r eval=FALSE}
is.this <- function(x) {
  if (x == this) {
    return(F)
  } else {
    return(T)
  }
}
lapply(alist, function(x) is.this(x))
```
