---
output: rmarkdown::github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r echo=FALSE, message=FALSE}
knitr::opts_chunk$set(message=FALSE, comment="")
```

# Cyclomatic Complexity of R (or any) Code

Cyclomatic complexity is a measurement developed by Thomas McCabe in 1976 to determine the stability and level of confidence in a program. It measures the number of linearly-independent paths through a program module. Programs with lower Cyclomatic complexity are easier to understand and less risky to modify.

The way Cyclomatic complexity is computed is explained [here](https://www.bbau.ac.in/dept/CS/TM/Cyclomatic.pdf)

The recommended complexity value is 10 or below, though in some circumstances, complexity as high as 15 is considered acceptable.  


# cyclocomp R package

Functions that do not have any conditions or branches has complexity of 1
```{r, warning=FALSE}
library(cyclocomp)

cyclocomp(
  
  
  my_summary <- function(x, y){
    ave.x <- mean(x, na.rm=T)
    ave.y <- mean(y, na.rm=T)
    xy.cor <- cor(x,y)
    c(ave_x = ave.x, ave_y = ave.y, xy_cor=xy.cor)
  }
  
)
```

An example with one `if()` statement
```{r}

cyclocomp(
  
  
  my_if <- function(x){
    if (any(is.na(x))) 0 else mean(x)
  }
  
)
```


More complex example with multiple conditions
```{r}

cyclocomp(
  
  
  my_if2 <- function(x){
    
    if (any(is.na(x)) | length(x) < 1){
      print("Error message")
      return(NA)
    } else {
      res <- mean(x)
    }
    
    if ( res == 0 ) print("warning") else return(res)
  }
  
)
```

Every loop adds additional complexity to the code as there is a condition 
that needs to be checked while loop is executing

```{r}

cyclocomp(
  
  my_for <- function(x){
    
    for (i in seq_along(x)){
      x[i] <- x[i] * i + i
    }
    
  }
  
)
```

# Using cyclocomp R package for R packages


```{r}

cyclocomp_package("cyclocomp")
```
If you are in the R project where you develop an R package, you can compute the 
project complexity using `cyclocomp_package_dir()` function:


```{r, eval=FALSE}

cyclocomp_package_dir(path = ".")
```


